---
title: "Non Linear Models"
subtitle: "Bayesian approach"
institution : Wits
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "sidebartab"
includes:
   in_header: header.txt
date: "Friday, April 7, 2017"
framework   : io2012

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
x = c( 1.0,  1.5,  1.5,  1.5, 2.5,4.0,  5.0,  5.0,  7.0,
       8.0,  8.5,  9.0,  9.5, 9.5,10.0, 12.0, 12.0, 13.0,
       13.0, 14.5, 15.5, 15.5, 16.5,17.0, 22.5, 29.0, 31.5)
Y = c(1.80, 1.85, 1.87, 1.77, 2.02,2.27, 2.15, 2.26, 2.47,
      2.19, 2.26, 2.40, 2.39, 2.41,2.50, 2.32, 2.32, 2.43,
      2.47, 2.56, 2.65, 2.47, 2.64,2.56, 2.70, 2.72, 2.57)

dataSet <- data.frame(age=x,length=Y)
```

## Non linear vs Linear approach

 - the linear mean structure is:  $Y_i=x_i^{\prime}\beta+\epsilon_i$
    - the generic form:  $Y_i=g(x_i, \beta) + \epsilon_i$ for a known function g
 - **Lets consider a non linear mean structure**
    - The idea is to model non transformed data

## Non transformed data
- The data are length and age measurements for 27 captured dugongs (sea cows). 
    - **Carlin and Gelfand (1991)** model this data using a nonlinear growth
 
***
```{r pressure}
plot(x,Y , ylab="Length", xlab="Age", main="Growth of dugongs")
lines(lowess(x,Y), col="blue")
```


## Non-linear Dugong growth model


$$
Y_i=\alpha - \beta*\gamma^{x_i}+  \epsilon_i, i=1,2,.....,n
$$

  - Where $\alpha>0$ , $\beta>0$ , $0\leq \gamma\leq1$ and as usual $\epsilon{_i}\sim N(0,\sigma^2)$ 
  -  $\alpha$ corresponds to the average length of a fully grown dugong
  -  $\alpha - \beta$ length of a dugong at birth and $\gamma$ determines the growth rate

## Sampling approach: why?
 - The nonlinearity of the model eliminates any hope for a closed form full conditional for $\gamma$
 - **Sampling** is the best approach - types of sampling?


## get data and code here?

 https://goo.gl/d5pbBo

## OpenBugs Model ..
  - We run three parallel Gibbs sampling chains of 20,000 iterations each
    following a 1000-iteration burn-in
  - Obtain posterior density estimates and autocorrelation plots for $\alpha>0$ , $\beta>0$ , $\gamma$ and $\sigma$

  - **Investigate the bivariate posterior of $(\alpha,\gamma)$ using the Correlation tool on the inference menu**

## Winbugs Code
```{r,  eval=FALSE, echo=TRUE}
model
	{
		for( i in 1 : N ) {
			Y[i] ~ dnorm(mu[i], tau)
			mu[i] <- alpha - beta * pow(gamma,x[i])			
		}
		alpha ~ dflat()
		beta ~ dflat()
		gamma ~ dunif(0.5, 1.0)
  U1 <- log(alpha);
  U2 <- log(beta);
  U3 <- logit(gamma);
  tau <- 1/(sigma*sigma)  
  sigma ~ dunif(0.01, 100)
	}

```


***
```{r,  eval=FALSE,echo=TRUE}
#data
list(
x = c( 1.0,  1.5,  1.5,  1.5, 2.5,   4.0,  5.0,  5.0,  7.0,
	8.0,  8.5,  9.0,  9.5, 9.5,  10.0, 12.0, 12.0, 13.0,
13.0, 14.5, 15.5, 15.5, 16.5, 17.0, 22.5, 29.0, 31.5),
Y = c(1.80, 1.85, 1.87, 1.77, 2.02, 2.27, 2.15, 2.26, 2.47,
	2.19, 2.26, 2.40, 2.39, 2.41, 2.50, 2.32, 2.32, 2.43,
2.47, 2.56, 2.65, 2.47, 2.64, 2.56, 2.70, 2.72, 2.57), N = 27)
#initial values
list(alpha = 1, beta = 1, sigma = 1, gamma = 0.9)
list(alpha = 10, beta = 10, sigma = 10, gamma = 0.7)
list(alpha = 100, beta = 100, sigma = 100, gamma = 0.5)
```


## Stata Implementation
```{r , eval=FALSE, echo=TRUE}
insheet using "Dugongs.csv" , clear
*****model1
bayesmh length ,dots(1000) ///
rseed(12345) saving(nl_growth_model1, replace) ///
mcmcsize(25000) burnin(5000) thinning(9) ///
likelihood(normal({var})) ///
prior({length:_cons}, density({alpha}-{beta}*{gamma}^age)) ///
prior({alpha}, flat)  ///
prior({beta}, flat) ///
prior({gamma}, uniform(0.5,1.0)) ///
block({beta} {alpha} {gamma} ) blocksummary ///
prior({var},igamma(0.001, 0.001))
***diagnostics graph
bayesstats ess _all
bayesgraph diagnostics _all
bayesgraph matrix _all

```


***
```{r, eval=FALSE , echo=TRUE}
****model2 - with more ierations compare with the bugs output
bayesmh length ,dots(1000) ///
rseed(2468) saving(nl_growth_model2, replace) ///
mcmcsize(100000) burnin(5000) thinning(9) ///
likelihood(normal({var})) ///
prior({length:_cons}, density({alpha}-{beta}*{gamma}^age)) ///
prior({alpha}, normal(0,1000))  ///
prior({beta},igamma(0.001,0.001)) ///
prior({gamma}, beta(0.5,1.0)) ///
block({beta} {alpha} {gamma} ) blocksummary ///
prior({var},igamma(0.001, 0.001))
**daignostics graphs
bayesstats ess _all
bayesgraph diagnostics _all
bayesgraph matrix _all

```


## Some  intro
https://youtu.be/3OJEae7Qb_o?list=PLTn3eOV1DiQi80T3K7vrB_7cXYaLNb-Y-


